---
title: "Lab Week 8"
author: "Jasmine Childress"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the packages
```{r}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)

```


### Column graph of Texas oil spills


```{r}

oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>%
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>% 
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x = Year, y = Loss))+
  geom_col()

```


### Leaflet plot of spill locations in TX in 2016
```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326) # making the data frame have "simple features" (sf), so that coordinates are sticky - they remain with data points

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers() # we only have one variable - net loss - so we don't need to specify anything in this command

```



### Make a tmap plot with the Texas state shapefile
- sometimes we can't use leaflet maps and need a shapefile of what we want to map
```{r}

states <- st_read(dsn = ".", layer = "states") # loading all the states files in my current working directory

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326)  #transforming the data here to maintain the same coordinate system


#plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)  # allo the locations in texas during 2016 where there were oil spills

```

###Convert the data to spatial points patterns (combination of the point data and the bounding window)

Answering the question: Are these patterns of oil spill completely spatially random (csr)?

```{r}

spills_sp <- as(oil_sf, "Spatial") #converting this data back from simple features to a spatial data fram
spill_ppp <- as(spills_sp, "ppp") # converting this to a point pattern analysis

tx_sp <- as(tex_border, "Spatial")

tx_owin <- as(tx_sp, "owin")     # texas outer window, what we will use to bound our data

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin) # rejects any points in space that aren't within the bounds of the window (if looking at the above map, there are a couple of obvious points outside the texas border)
```

### A density plot:
```{r}

plot(density(all_ppp, sigma = 0.4)) # density plots are easy to make, but can change the story of your data. Changing sigma alters the visual representation of the plot.

```


### Quadrat test for spatial evenness
- Are oil spills evenly distributed throughout Texas for 2016?

```{r}

oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) # horizontally (nx) and vertically (ny) there are 5 regions. Counts the number of events and figures out the intensity of each quadrat and performs a chi-squared test amongst the quadrats.
oil_qt
# p-value is very small indicating that we reject the null hypothesis that the data follows completely spatial randomness. The data is NOT evenly distributed. 

plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4)

# Looking at the upper left quadrat: the number of expected events would be 9 because it is a partial quadrat. The upper left number (2) is the actual number of events. On the bottom, there is a standardized measure of how different it is from the expected in terms of standard deviations (-2.3). For full quadrats, the expected number of counts is 14.5 if the data were truly even. 

```

### G-Function for nearest neighbor analysis
```{r}

r <- seq(0,1, by = 0.01)

# based on a Poison distribution, what would this data look like if it was truly completely spatial randomness?

oil_gfun <- envelope(all_ppp, fun = Gest, r = r, nsim = 100) # looking for one point's single nearest neighbor, which is why this runs quickly.

ggplot(oil_gfun, aes(x = r, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red") # This graphs tells us that our observed data has a higher proportion of point pairss with nearest neighbors of shorter distances compared to modeled data that was completely spatially random.

```


















